---
title: "Misinformation Analysis"
author: "Daniel Allington"
date: "2025-03-06"
output:
  word_document: default
  html_document: default
---

```{r setup}
library(dplyr)
library(readr)
library(ggplot2)

options(scipen = 999)

misinfo_d <- read_csv(
  '../data/life-under-lockdown-wave2-extract.csv',
  show_col_types = FALSE)

knowledge_levels <- c(
  'Nothing at all',
  'Not very much',
  'A fair amount',
  'A great deal'
)

misinfo_d <- misinfo_d %>%
  mutate(
    know_facebook = know_facebook %>%
      ordered(
        levels = knowledge_levels
      )
  )
```

# Hypotheses
H1 We expect a relationship between gender and number of misinformation beliefs

H2 We expect a relationship between age and number of misinformation beliefs

H3 We expect a relationship between relying on Facebook for knowledge and number of misinformation beliefs

# Descriptive statistics
```{r visualise-age}
misinfo_d %>%
  ggplot(
    aes(x = age)
  ) +
  geom_histogram(binwidth = 5)
```

```{r visualise-gender}
misinfo_d %>%
  ggplot(
    aes(x = gender)
  ) +
  geom_bar()
```

```{r visualise-facebook-knowledge}
misinfo_d %>%
  ggplot(
    aes(x = know_facebook)
  ) +
  geom_bar()
```

```{r visualise-beliefs}
misinfo_d %>%
  ggplot(
    aes(x = beliefs)
  ) +
  geom_histogram(binwidth = 1)
```

# Findings
## Number of misinformation beliefs versus gender
```{r gender-beliefs-table}
misinfo_d %>%
  group_by(gender) %>%
  summarise(
    n = n(),
    M = mean(beliefs),
    Mdn = median(beliefs),
    SD = sd(beliefs)
  )
```

```{r gender-beliefs-test}
misinfo_d %>%
  filter(gender != 'In another way') %>%
  t.test(
    beliefs ~ gender,
    data = .
  )
```

In the sample, men had more misinformation beliefs (M = 0.73, SD = 1.14) than women (M = 0.61, SD = 1.05), but this was not statistically significant, t(494.31) = -1.201, p = 0.230.

## Number of misinformation beliefs versus age
```{r age-beliefs-visualisation}
misinfo_d %>%
  ggplot(
    aes(x = age, y = beliefs)
  ) +
  geom_jitter()
```

```{r age-beliefs-correlation}
misinfo_d %>%
  cor.test(
    ~ age + beliefs,
    data = .,
    method = 'pearson'
  )
```

In the sample, there is a statistically significant moderate negative correlation between age and number of misinformation beliefs, r(498) = -.24, 95% CI [-.32, -.16], p < .001.

## Number of misinformation beliefs versus reliance on Facebook for knowledge
```{r facebook-beliefs-visualisation}
misinfo_d %>%
  ggplot(
    aes(x = know_facebook, y = beliefs)
  ) +
  geom_boxplot()
```

```{r facebook-beliefs-correlation}
misinfo_d %>%
  mutate(
    know_facebook = as.numeric(know_facebook)
  ) %>%
  cor.test(
    ~ know_facebook + beliefs,
    data = .,
    method = 'spearman',
    exact = FALSE
  )
```

In the sample, there is a statistically significant moderate positive correlation between relying on Facebook for knowledge about coronavirus and number of misinformation beliefs, rs(498) = .23, p < .001.



## Linear model
```{r dummy-coding}
model_d <- misinfo_d %>%
  mutate(
    Female = as.numeric(gender == 'Female'),
    High_Facebook = as.numeric(
      know_facebook %in% c(
        'A fair amount', 'A great deal')
      )
  )
```

```{r age-model}
model_d %>%
  lm(
    beliefs ~ age,
    data = .
  ) %>%
  summary
```

```{r facebook-model}
model_d %>%
  lm(
    beliefs ~ High_Facebook,
    data = .
  ) %>%
  summary
```

```{r gender-model}
model_d %>%
  lm(
    beliefs ~ Female,
    data = .
  ) %>%
  summary
```

```{r full-model}
model_d %>%
  lm(
    beliefs ~ age + High_Facebook + Female,
    data = .
  ) %>%
  summary
```



beliefs = 1.22 + -0.01 * age + 0.58 * High_Facebook + -0.17 * Female

Female Facebook user: beliefs = 1.22 - age / 100 + 0.59 - 0.17
Male Facebook user: beliefs = 1.22 - age / 100 + 0.59
Female Facebook non-user: beliefs = 1.22 - age / 100 - 0.17
Male Facebook non-user: beliefs = 1.22 - age / 100

